---
title: Data Extraction
author: Matthew Leonawicz
output:
  html_document:
    toc: true
    theme: united
    highlight: zenburn
    keep_md: true
---

```{r knitr_setup, echo=FALSE}
opts_chunk$set(cache=FALSE, echo=FALSE, eval=TRUE, tidy=TRUE, message=FALSE, warning=FALSE)
read_chunk("../../code/extract_eval.R")
```

### Next steps

Combining sampling and data reduction methods while using the most efficient **R** functions can be particularly useful when processing large numbers of high-resolution geotiff raster layers.
One thing I already do when extracting from many files by shapefile is I avoid extracting by shape more than once.
I do it one time to obtain the corresponding raster layer cell indices.
Then on all subsequent maps I extract by cell indices which is notably faster.
Ultimately, there is much more room for speed improvements in terms of efficient use of statistics than in strictly programmatic corner-cutting.

The plots below benchmark different sample mean computations.
Comparisons involve the sample mean of the entire data set and do not involve the main approach outlined above which focuses on efficiency gains by taking the mean of a smaller, representative sample.
This provides some insight into how it is beneficial nonetheless to considering the right programmatic approach in conjunction with statistical efficiencies.

```{r benchmarks1}
```

```{r benchmarks2}
```
